---
title: "Introduction to 2018 Statistics Methods Forum Data Challenge"
author: "Eric Polley"
date: "June 27th, 2018"
output: beamer_presentation
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```
## Introduction

The focus this year is the estimation and evaluation of a prognostic risk score with a right censored outcome of interest.

The dataset is from a clinical trial in Non-Small Cell lung cancer with N=398 patients. The patients were randomly split into training (N=300) and a blinded test set (N=98).

Details for the Data Challenge are available on Github: \url{https://github.com/ecpolley/Data_Challenge_2018}

## Outline

The data challenge will use the next 2 Statistical Methods Forum meetings

* June 27th: Introduction to the data challenge
* July 25th: Group Discussion and Q&A
* August 20th 5:00pm local: Team submission deadline
* August 22nd: Final Results presentation

## Team Science

* Participants are encouraged to work in teams\newline (N $\in$ (1, 2, \ldots, 10))
* Opportunity to learn from each other and work with people outside usual team
* Data is publicly available, so is available outside Mayo 
* If you would like help forming a team, email Eric Polley or Kristin Mara
* Teams are responsible for creating a team name, and may submit up to 3 estimates, with the last submission being the official one
* If you are participating, please let us know in case we have any Data Challenge announcements

## Overview of Dataset

* Two datasets will be provided for the training and test sets
* Clinical dataset with baseline variables and outcome
* Each patient had a baseline Lung CT scan, single slice compiled in an (512, 512, N) Array

## Overview of Dataset

```{r loadData, echo = TRUE}
# link to data on GitHub page if not available
if(file.exists("Training_clinical.csv")) {
  dat <- read.csv("Training_clinical.csv")
} else {
  urlfile <- "https://raw.githubusercontent.com/ecpolley/
    Data_Challenge_2018/master/Training_clinical.csv"
  download.file(urlfile, destfile = "Training_clinical.csv")
  dat <- read.csv("Data.csv")
}
dim(dat)
```

## setup

```{r loadPackages, echo = TRUE}
library(arsenal)
library(survival)
library(survminer)
```


## Overview of Dataset
\tiny
```{r sumData, echo = FALSE, results = 'asis'}
tab1 <- tableby( ~ age + as.factor(Clinical.T.Stage) + as.factor(Clinical.N.Stage) + Clinical.M.Stage + Overall.Stage + Histology + gender, data = dat)
summary(tab1)
```

## Overview of Dataset
```{r plotData, echo = FALSE}
fit <- survfit(Surv(Survival.time, deadstatus.event) ~ 1, data = dat)
ggsurvplot(fit, data = dat, risk.table = TRUE)

```

## Overview of Dataset
```{r loadImg, echo = TRUE}
load("DataChallengeDataTrain_array.RData")
dim(IMAGES_array_train)
```

## Overview of Dataset
```{r loadImg24, echo = TRUE}
image(IMAGES_array_train[, , 4])
```

## Goal

* The primary goal is to develop a prognostic risk score
* How to evaluate on the held out set?

## Evaluation

Primary goal:

* Each each patient in test set, provide predicted risk score
* Evaluate discrimination by estimating concordance with observed survival times
* Each team can email me (\url{Polley.Eric@Mayo.edu}) with text file including patient ID and predicted risk score


## Evaluation

Secondary goal:

* Evaluate calibration of predicted probability of survival at specific time points
* For each patient in test set, provide predicted probability of survival at 1, 2, and 3 years post treatment
* 365, 730, and 1095 days
* For each time point, split test data into quintiles based on predicted probability. Compare to Kaplan-Meier estimate.


## Questions?
